% !TEX root = FinalProject.tex

\chapter{Problem Statement}

For this project, our team decided to apply Natural Computing to one of the most
popular fields of general computing: games. Rather than create a game that uses
ideas from natural computing to enhance some existing element, however, we opted to
create a game entirely based on some concepts from Natural Computing.

This game will allow players to control a species of animal on an island, to
compete with other players' animals and try to become the dominant species. The
player will have some say in what sort of animal they want to create, by setting
up the parameters via which their animals evolve. Beyond that, they simply sit back
and watch as their population grows and shifts, exploring the island and fighting
the other players' creations.

The purpose of this is really two-fold. Primarily, we can't help ourselves - every
time we learn a new concept, we tend to imagine some fantastic way to use that
concept; when it comes to natural computing, these concepts lend themselves very
easily to games. \\
Beyond that, a game based on Natural Computing would fit a particular niche. The
ideas of natural computing are fairly user-independent, but there is an entire
category of simulation games that work this way, letting the user set up initial
parameters and merely updating them on their progress.\\
Moreover, this game would provide a sort of educational experience. By allowing
players to modify evolutionary parameters and watch the results unfold in front of
them, we are giving them the opportunity to reason through why these changes
occured and consider what actions they can take to nudge the computer's search
toward their goal of island domination.\\
Finally, this project will be a great opportunity for the team to apply what
they have learned in the course thus far. This project will doubtlessly draw on
many different aspects of Natural Computing, and this will force the team to
consider how these different aspects can work together, and to analyze the results
they get. When making a game, it is important that the system be balanced and fair;
and in Natural Computing, a fair system is only one point in a massive grid of
possibilities.
